---
title: "clusteranimaltimeseries"
author: "Kate Sheridan"
date: "9/11/2021"
output: html_document
---

```{r setup, include=FALSE}
library(stringr)
library(cluster)
library(nomclust) 
library(ggplot2)
#library(reshape2)
library(ggthemes)
library(fpc)
library(gridExtra)
library(tidyverse)
library(here)
library(Rtsne)
library(gclus)
library(vegan)
library(ggdendro)
library(tsne)
library(ggfortify)
library(dplyr)
```

## load in data

```{r load-in}
#load in imputed data
animal_imp2 <- read.csv(file = here('data', 'clean','20210806_animal_imputed.csv'))
algae_imp2 <- read.csv(file = here('data', 'clean','20210806_algae_imputed.csv'))

#remove x column
animal_imp2 <- animal_imp2[,2:ncol(animal_imp2),]
algae_imp2 <- algae_imp2[,2:ncol(algae_imp2),]

#make factors
animal_imp2 <- animal_imp2 %>% 
  mutate(across(.fns = ~ as.factor(.)))
algae_imp2 <- algae_imp2 %>% 
  mutate(across(.fns = ~ as.factor(.)))


# raw if needed
#be sure to use na.strings or imputation will fail!
trait_raw <- read.csv(here("data", '20210730_ShoalsFunctionalTraits.csv'), 
                      na.strings = c("","NA", 'N/A'))
```

## load in timeseries

```{r}
#uncomment the one you need
sp_list <- read.csv(here('data', 'fielddata', 'sp_1982r.csv'))
#sp_list <- read.csv(here('data', 'fielddata', 'sp_1996r.csv'))
#sp_list <- read.csv(here('data', 'fielddata', 'sp_2011r.csv'))

algae_imp2 <- algae_imp2 %>%
  subset(species %in% sp_list$value)

animal_imp2 <- animal_imp2 %>%
  subset(species %in% sp_list$value)
```

## load in custom functions + clustering info

```{r}
#set up evaluations

##list of the possible internal validation measures
nomclustEvals <- c("PSFM", "PSFE", "BIC", "AIC")

#old version
#nomclustEvals <- c("WCM", "WCE", "PSTau", "PSU", "PSFM", "PSFE")
#List of the possible linkage methods
clusterMethod <- c("average", "complete", "single")


#Colours for the different distance matrices
group.colours <- c("Goodall" = "black",
                   "Goodall_2" = "#CCCCCC",
                   "Goodall_3" = "#999999",
                   "Goodall_4" = "#666666",
                   "Eskin"="blue",
                   "IOF"="orange", 
                   "OF" = "#D55E00",
                   "Lin"="purple",
                   "Lin_1" = "#9999CC",
                   "Simple matching"="darkgreen")

subDist <- c("Goodall","Eskin", "IOF", "Lin", "Simple matching")

#relevant functions from Ladds 2018
fit_nomclust <- function(dataSET, distanceMethod,clusterMethod, cmax)
                        {nomclust(data = dataSET, measure = distanceMethod, 
                                           method = clusterMethod,
                                  clu.high = cmax)}


plotEvals <- function(dat, measure){
  ggplot(dat, 
         aes(x = cluster, y = value, color = distance))+
    geom_point()+geom_line()+
    theme_base()+
    ylab(measure)+
    xlab("Clusters")+
    scale_x_discrete(limits = c(2:9))+
    #scale_y_continuous(limits = c(lowylim,highylim))+
    scale_color_manual(values = group.colours)+
    theme(panel.grid.minor.x	= element_line(color = "grey"),
          panel.grid.major.x	= element_line(color = "grey"),
          panel.grid.major.y	= element_line(color = "grey"),
          legend.title = element_blank())
}





groupfits <- function(final_fit,distMeasure,clusterMethod){
  dataSET$groups <- final_fit$mem
  cor.matrix <- sapply(final_fit$mem, function(x) as.matrix(dist(x,method = "manhattan")))
  
  for(k in 1:8) {
    m2 <- melt(matrix(cor.matrix[,k],nrow = nrow(dataSET)))[melt(upper.tri(matrix(cor.matrix[,k],nrow = nrow(dataSET))))$value,]
    m2$value<-ifelse(m2$value==0,"Match","NoMatch")
    names(m2) <- c("c1", "c2", paste0(distMeasure,"-",clusterMethod,"_",k+1))
    m2$c1<-factor(m2$c1,labels = namesComm[1:nrow(dataSET)-1])
    m2$c2<-factor(m2$c2,labels = namesComm[2:nrow(dataSET)])
    comparison <- merge(comparison,m2,by=c("c1","c2"))
  }
  return(comparison)
}
```


```{r dissimilarity matrix legendre}
"coldiss" <- function(D, nc = 4, byrank = TRUE, diag = FALSE)
{
	require(gclus)

	if (max(D)>1) D <- D/max(D)

	if (byrank) {
		spe.color = dmat.color(1-D, cm.colors(nc))
	}
	else {
		spe.color = dmat.color(1-D, byrank=FALSE, cm.colors(nc))
	}

	spe.o = order.single(1-D)
	speo.color = spe.color[spe.o,spe.o]
	
	op = par(mfrow=c(1,2), pty="s")

	if (diag) {
		plotcolors(spe.color, rlabels=attributes(D)$Labels, 
			main="Dissimilarity Matrix", 
			dlabels=attributes(D)$Labels)
		plotcolors(speo.color, rlabels=attributes(D)$Labels[spe.o], 
			main="Ordered Dissimilarity Matrix", 
			dlabels=attributes(D)$Labels[spe.o])
	}
	else {
		plotcolors(spe.color, rlabels=attributes(D)$Labels, 
			main="Dissimilarity Matrix")
		plotcolors(speo.color, rlabels=attributes(D)$Labels[spe.o], 
			main="Ordered Dissimilarity Matrix")
	}

	par(op)
}
```


# Animals

setup
```{r animal matrix}
##Cluster using nomclust

##Make a NULL dataset for the output from "nomclust"
#This makes a long dataset with all possible comparisons
animal_combdf <- combn(animal_imp2$species,2)
animal_comparison<-data.frame(animal_combdf[1,], animal_combdf[2,])
colnames(animal_comparison) <- c("c1", "c2")

#Select the variables to use from the imputed data
animal_dataSET <- animal_imp2[,c(1:14)]
animal_spnames <- animal_imp2[,"species"]

#other variables for matrix
evaluation <- NULL
cmax = 10


animal_dataSET2 <- data.frame(animal_dataSET[,-1], row.names = animal_dataSET[,1])
```

generate all options with loop from ladds
```{r animal loop}

dataSET <- animal_dataSET2
animal_eval <- {}

# #heirachical clustering#
 for(i in 1:length(clusterMethod))  {
# 
#   ##---Goodall 1
# 
#   #fit the model
   good_fit<-fit_nomclust(dataSET, 'good1',clusterMethod[i], cmax)
   #comparison<-groupfits(good_fit,"good1",clusterMethod[i])
# 
#   #Extract the validation metrics
   good_eval <- dplyr::bind_rows(good_fit$eval)
   good_eval$distance <- "Goodall"
   good_eval$method <-  clusterMethod[i]
# 
#   ##---Goodall 2
# 
   good2_fit<-fit_nomclust(dataSET, 'good2',clusterMethod[i], cmax)
   #comparison<-groupfits(good2_fit,"good2",clusterMethod[i])
# 
#   #Extract the validation metrics
  good2_eval <- dplyr::bind_rows(good2_fit$eval)
  good2_eval$distance <- "Goodall_3"
  good2_eval$method <-  clusterMethod[i]
# 
#   ##---Goodall 3
   good3_fit<-fit_nomclust(dataSET, 'good3',clusterMethod[i], cmax)
#   comparison<-groupfits(good2_fit,"good3",clusterMethod[i])
# 
#   #Extract the validation metrics
   good3_eval <- dplyr::bind_rows(good3_fit$eval)
   good3_eval$distance <- "Goodall_3"
   good3_eval$method <-  clusterMethod[i]
# 
# 
#   ##---Goodall 4
# 
   good4_fit<-fit_nomclust(dataSET, 'good4',clusterMethod[i], cmax)
#   comparison<-groupfits(good4_fit,"good2",clusterMethod[i])
# 
#   #Extract the validation metrics
   good4_eval <- dplyr::bind_rows(good4_fit$eval)
   good4_eval$distance <- "Goodall_4"
   good4_eval$method <-  clusterMethod[i]
# 
#   ##---Eskin
# 
   eskin_fit <- fit_nomclust(dataSET, 'eskin',clusterMethod[i], cmax)
#   comparison<-groupfits(eskin_fit,"eskin",clusterMethod[i])
# 
   eskin_eval <- dplyr::bind_rows(eskin_fit$eval)
   eskin_eval$distance <- "Eskin"
   eskin_eval$method <- clusterMethod[i]
# 
# 
#   #--IOF
# 
   iof_fit <- fit_nomclust(dataSET, 'iof', clusterMethod[i], cmax)
#   comparison<-groupfits(iof_fit,"iof",clusterMethod[i])
# 
   iof_eval <- dplyr::bind_rows(iof_fit$eval)
   iof_eval$distance <- "IOF"
   iof_eval$method <- clusterMethod[i]
# 
#   #--OF
# 
   of_fit <- fit_nomclust(dataSET, 'of',clusterMethod[i], cmax)
#   comparison<-groupfits(of_fit,"of",clusterMethod[i])
# 
   of_eval <- dplyr::bind_rows(of_fit$eval)
   of_eval$distance <- "OF"
   of_eval$method <- clusterMethod[i]
# 
# 
#   #--Lin
# 
   lin_fit <- fit_nomclust(dataSET, 'lin', clusterMethod[i], cmax)
#   comparison<-groupfits(lin_fit,"lin",clusterMethod[i])
# 
   lin_eval <- dplyr::bind_rows(lin_fit$eval)
   lin_eval$distance <- "Lin"
   lin_eval$method <- clusterMethod[i]
# 
# 
#   #--Lin1
# 
   lin1_fit <- fit_nomclust(dataSET, 'lin1',clusterMethod[i], cmax)
#   comparison<-groupfits(lin1_fit,"lin",clusterMethod[i])
# 
   lin1_eval <- dplyr::bind_rows(lin1_fit$eval)
   lin1_eval$distance <- "Lin_1"
   lin1_eval$method <- clusterMethod[i]
# 
# 
#   #--Simple matching
# 
   sm_fit <- fit_nomclust(dataSET, 'sm', clusterMethod[i], cmax)
#   comparison<-groupfits(sm_fit,"sm",clusterMethod[i])
# 
   sm_eval <- dplyr::bind_rows(sm_fit$eval)
   sm_eval$distance <- "Simple matching"
   sm_eval$method <- clusterMethod[i]
# 
   animal_eval <- rbind(animal_eval, good_eval, good2_eval,
                       good3_eval, good4_eval,eskin_eval,
                       iof_eval, of_eval, lin_eval,
                       lin1_eval, sm_eval
)
# 
 }


#save rdata if needed / don't want to rerun
#save(animal_eval, file = here('output', 'cluster_ks', 'animal_evaluation_2-25.RData'))

```

make long and scale for evaluation steps
```{r}
#make long
animal_eval <- animal_eval %>%
    na_if('NaN') %>%
  select(!(c(WCM, WCE, BK))) %>%
  pivot_longer(cols = c(PSFM:AIC)) %>%
  dplyr::rename(cluster = names) %>%
  subset(!(is.na(value)))

#make factors
animal_eval <- animal_eval %>% 
  mutate(across(cluster:name, .fns = ~ as.factor(.)))

#change clusters into numbers!!!!
animal_eval$cluster <- str_replace_all(c(animal_eval$cluster), "clu\\_", "")
animal_eval$cluster <- as.numeric(animal_eval$cluster)

#scale
animal_eval$value_scaled <- animal_eval$value/max(animal_eval$value)
```

## validation/evaluation for each method to pick number of clusters and matrix type 
```{r animal-validation}
##------------ VALIDATION

## Validation from the nomclust package internal measures. 
#Sulc (2016) thesis; pg. 31

nomclustEvals <- c("PSFM", "PSFE", "BIC", "AIC")

evaluation <- animal_eval

###PSFE - higher value indicates better grouping.
EvalData <- evaluation[evaluation$name=='PSFE'&!evaluation$cluster==1,]

##If you want to subset further
#EvalData <- EvalData[!EvalData$cluster==2&EvalData$distance %in% subDist,]

EvalPlotPSFE <- ggplot(EvalData, 
                       aes(x = cluster, y = value, color = factor(distance)))+
  geom_point()+geom_line()+
  theme_base()+
  ylab("PSFE")+
  xlab("Clusters")+
  scale_x_discrete(limits = c(3:25))+
  facet_grid(rows = vars(method)) +
  #scale_y_continuous(limits = c(lowylim,highylim))+
  scale_color_manual(values = group.colours)+
  theme(panel.grid.minor.x	= element_line(color = "grey"),
        panel.grid.major.x	= element_line(color = "grey"),
        panel.grid.major.y	= element_line(color = "grey"),
        #axis.text.x = element_blank(),
        legend.title = element_blank(),
        legend.background = element_rect(fill = "white", colour = NA),
        plot.background = element_blank())


###PSFM - higher value indicates better grouping.
EvalData <- evaluation[evaluation$name=='PSFM'&!evaluation$cluster==1,]

#EvalData <- EvalData[!EvalData$cluster==2&EvalData$distance %in% subDist,]
#EvalData$distance <- droplevels(EvalData$distance)
EvalPlotPSFM <- ggplot(EvalData, 
                       aes(x = cluster, y = value, color = factor(distance)))+
  geom_point()+geom_line()+
  theme_base()+
  ylab("PSFM")+
  xlab("Clusters")+
  scale_x_discrete(limits = c(3:25))+
  facet_grid(rows = vars(method)) +
  #scale_y_continuous(limits = c(lowylim,highylim))+
  scale_color_manual(values = group.colours)+
  theme(panel.grid.minor.x	= element_line(color = "grey"),
        panel.grid.major.x	= element_line(color = "grey"),
        panel.grid.major.y	= element_line(color = "grey"),
        #axis.text.x = element_blank(),
        legend.title = element_blank(), 
        plot.background = element_blank())

###AIC -XX value indicates better grouping.
EvalData <- evaluation[evaluation$name=='AIC'&!evaluation$cluster==1,]

#EvalData <- EvalData[!EvalData$cluster==2&EvalData$distance %in% subDist,]
#EvalData$distance <- droplevels(EvalData$distance)
EvalPlotAIC <- ggplot(EvalData, 
                       aes(x = cluster, y = value, color = factor(distance)))+
  geom_point()+geom_line()+
  theme_base()+
  ylab("AIC")+
  xlab("Clusters")+
  scale_x_discrete(limits = c(3:25))+
  facet_grid(rows = vars(method)) +
  #scale_y_continuous(limits = c(lowylim,highylim))+
  scale_color_manual(values = group.colours)+
  theme(panel.grid.minor.x	= element_line(color = "grey"),
        panel.grid.major.x	= element_line(color = "grey"),
        panel.grid.major.y	= element_line(color = "grey"),
        #axis.text.x = element_blank(),
        legend.title = element_blank(), 
        plot.background = element_blank())


###AIC -XX value indicates better grouping.
EvalData <- evaluation[evaluation$name=='BIC'&!evaluation$cluster==1,]

#EvalData <- EvalData[!EvalData$cluster==2&EvalData$distance %in% subDist,]
#EvalData$distance <- droplevels(EvalData$distance)
EvalPlotBIC <- ggplot(EvalData, 
                       aes(x = cluster, y = value, color = factor(distance)))+
  geom_point()+geom_line()+
  theme_base()+
  ylab("BIC")+
  xlab("Clusters")+
  scale_x_discrete(limits = c(3:25))+
  facet_grid(rows = vars(method)) +
  #scale_y_continuous(limits = c(lowylim,highylim))+
  scale_color_manual(values = group.colours)+
  theme(panel.grid.minor.x	= element_line(color = "grey"),
        panel.grid.major.x	= element_line(color = "grey"),
        panel.grid.major.y	= element_line(color = "grey"),
        #axis.text.x = element_blank(),
        legend.title = element_blank(), 
        plot.background = element_blank())

```
## save plots, be sure to change name / location if needed
```{r save-plots}

#Plot and export the different measures  
for(i in 1:length(nomclustEvals)){
  EvalData <- evaluation[evaluation$name==nomclustEvals[i],]
  EvalPlot <- plotEvals(EvalData, nomclustEvals[i])
  ggsave(EvalPlot, device = "pdf",
         filename = here('output', 'cluster_ks',  paste0("animal_", nomclustEvals[i],"_plot-2011r.pdf")))
}

ggsave(EvalPlotPSFE, filename = here('output', 'cluster_ks', 'Animal_EvalPlotPSFE-2011r.pdf'),
       width = 9, height = 4)

ggsave(EvalPlotPSFM, filename = here('output', 'cluster_ks', 'Animal_EvalPlotPSFM-2011r.pdf'), 
       width = 9, height = 4)

ggsave(EvalPlotAIC, filename = here('output', 'cluster_ks', 'Animal_EvalPlotAIC-2011r.pdf'), 
       width = 9, height = 4)

ggsave(EvalPlotBIC, filename = here('output', 'cluster_ks', 'Animal_EvalPlotBIC-2011r.pdf'), 
       width = 9, height = 4)


#ggsave(here('output', 'cluster_ks', 'animal-cluster-evalplotaic-1982r.png'), plot = EvalPlotAIC)
```

```{r}
#pick the method you want for matrix based on evaluation plots above
animal_matrix <- good3(animal_dataSET2)


#legendre dissimilarity matrix
coldiss(animal_matrix) 

#mds
animalmds <- metaMDS(animal_matrix)

plot(animalmds)
coldiss(animal_matrix)
#save these as graphs for looking at later

animal_matrix2 <- fortify(animal_matrix)
```

## save dissimilarity matrix and the visualization
```{r}
#save(animal_matrix, file = here('data','animal_matrix-1982r.RData'))
write.csv(animal_matrix2, file = here('data', 'clean', 'animal_matrix-2011r.csv'))

png(filename = here('output', 'cluster_ks', 'animal-dissimilarity-2011r.png'),
    width = 1000, height = 500,
    units = 'px')
coldiss(animal_matrix)
dev.off()
```

Here you will use the number of clusters identified above, as well as the method

```{r}
#define options for heirarchcal clustering 
methodsClust <- c("single", "complete", "average")

######Animals
animal_fit <- hclust(animal_matrix, method = methodsClust[3])  #change cluster method
animal_clust <- cutree(animal_fit, 4)  # change number of clusters

animal_dendr <- dendro_data(animal_fit, type="rectangle") 

animal_labs <- label(animal_dendr)
animal_labs$cluster <- as.factor(animal_clust)
  
animal_dendr_plot <- ggplot() +
  geom_segment(data=segment(animal_dendr), aes(x=x, y=y, xend=xend, yend=yend)) +
  geom_text(data=animal_labs, aes(x=x, y=y, label=label, hjust=0, color = cluster), size=2) +
  coord_flip() + scale_y_reverse(expand=c(0.1, 0), limits = c(.9,-0.1)) +
  theme_void()
  

animal_dendr_plot

ggsave(here('output', 'animal-cluster-dendro-1982r.png'), plot = last_plot())

# also make csv lists of animals in those years
animal_clust_l <- animal_labs %>%
  select(label, cluster)

write.csv(animal_clust_l, here('output', 'cluster_ks', 'animal-clusters-1982r.csv'))
```

## tsne

```{r}
## animals

animal_phyla <- trait_raw %>%
  subset(group %in% c('Invertebrate', 'Vertebrate')) %>%
  dplyr:::select(species, phylum)

animal_phyla <- animal_phyla %>%
  subset(species %in% sp_list$value)

tsne_obj_an <- Rtsne(animal_matrix, is_distance = TRUE, perplexity = 5)
plot(tsne_obj_an$Y)

tsne_plot_an <- data.frame(x = tsne_obj_an$Y[,1], y = tsne_obj_an$Y[,2], 
                         animal_dataSET2, animal_phyla, animal_clust_l)

tsne_animal <- ggplot(tsne_plot_an, aes(x=x, y=y)) + 
  #geom_point(aes(color = common_name_division_bgr)) +
  geom_point(aes(color = phylum)) +
  #geom_point(aes(color = cluster)) +
  theme_classic()

tsne_animal

ggsave(here('output', 'cluster_ks', 'animal-cluster-tsne_1982r-phy.png'), plot = last_plot())

```

